---
title: "Reproducible Research: Happy-Git"
author: "Philip Leftwich"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float: true
    highlight: pygments
    css: ../lab.css
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Why Git? 
Two reasons that we'll be most interested in (to start):

- It's a great way to safely store organized versions of a project

- It's ideal for data science collaborators who need to share & edit code together


# Some notes:
 
Repositories (repos) on GitHub are the same unit as an RStudio Project - it's a place where you can easily store all information/data/etc. related to whatever project you're working on.
 
When we create a Repository in GitHub and have it communicating with a Project in RStudio, then we can get (**pull**) information from GitHub to RStudio, or **push** information from RStudio to GitHub where it is safely stored and/or our collaborators can access it. It also keeps a *complete history* of updated versions that can be accessed/reviewed by you and your collaborators at any time, from anywhere, as long as you have the internet.

A couple of general tips:
 
- Pull frequently (if working with anyone else, and when you start working on a project again, even if on the same device)

- Commit/push in small, meaningful increments and with useful (searchable, descriptive) Commit messages

- The best way to deal with merge conflicts is to avoid creating merge conflicts. This is what happens when two people pull a version, work on it separately and then try to push back to the repo *at the same time* Communicate with collaborators so you’re not all working on the same piece of code at the same time.

# Learning outcomes

**PART 1.** Fork & Clone an existing repo

**PART 2.** Create your own repo and version controlled R project from scratch

**PART 3.** Use the GitHub Classroom to fork a project, organise folders, run a script 

**END**



# PART 1. Fork & clone an existing repo on GitHub, make edits, push back

a. Go to [github.com](https://github.com/) and log in (you probably are already)

b. In the Search bar, look for repo 

c. Click on the repo name, and check out the existing repo structure

d. FORK the repo

e. Press Clone/download, copy the URL, and create a new version controlled project in RStudio (follow along with video)


VIDEO


f. Open the some_cool_animals.Rmd document, and the html 

g. Add your name to the top of the document

h. BUT WAIT. We have forgotten to add a great image and facts about a very important species - Baby Yoda, including an image 

IMAGE

FACTS

i. Once you’ve added Grogu, knit the Rmd document to update the html

j. Stage, commit & push all files 

NOTES

k. On GitHub, refresh and see that files are updated. Cool! Now you’ve used something someone else has created, customized it, and saved your updated version. 

# PART 2. Create your own repo & version controlled R Project from scratch

a. Go back to your GitHub account

b. Click on the plus sign (upper right, by your profile pic/icon) to create a new repository

c. Name the repo bio-5023Y-happy-git-yourinitials (like Bio-5023Y-Happy-Git-pl), and select to initialize with a ReadMe

d. Edit the ReadMe (however you want - notice that markdown formatting works & you can see a preview) & commit

e. Clone to create a connected R Project in RStudio

f. Create a new R Markdown document

g. Attach the {tidyverse}, {DT}, and {plotly} packages in a hidden code chunk (include = FALSE)

h. Create an interactive plot of PalmerPenguins with {plotly}, showing the output but **not** the code or messages

EXAMPLE

i. Create an interactive table of XXX {DT}

j. Knit & save your .Rmd

k. Stage, commit & push back to GitHub


# PART 3. GitHub Classroom enabled R Projects with subfolders

BLURB ON CLASSROOM

a. Follow this [invite link]()

b. You will be invited to sign-in to Github (if not already) & to join the UEABIO organisation

c. Clone your assignment to work locally in RStudio

d. In your local project folder, create subfolders ‘data’ and ‘final_graphs’ (*Note* use the dir.create commands in the console)

e. Drop the file disease_burden.csv into the ‘data’ subfolder (*Note* unlike in Bash there is no base R command to move files so do this using the RStudio pane)

f. Open a new .R script

g. Attach the {tidyverse} package

h. Read in the disease_burden.csv data

i. Stage, commit & push at this point - notice that the empty folder ‘final_graphs’ doesn’t show up (won’t commit an empty folder)

j. Back in the script, HERE!!!!!! instructor to do some cleaning, wrangling & viz

k. Update your graph with direct labels (using annotate) and vertical or horizontal lines with geom_vline or geom_hline

l. Use ggsave() to write your graph to a .png in the ‘final_graphs’ subfolder

m. Save, stage, commit & push 

n. Check that changes are stored on GitHub

(*NOTE* this will be in your organisations rather than repos)

 Forking is disabled on these repos, we will use GitHub Classroom as a way to submit code for our next summative assignment. It allows us to have multiple forked repos all in one place, private to students but accessible to markers. 


```{r}

```
